Validation of Visual Statistical Inference, with Application to Linear Models

### 2012-03-04 paper_jasa.tex
tightening paper, restructured section on power considerations for lineup

@@

\section{Introduction} Any statistical analysis must include some statistical graphics. For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics. Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.

@@

\red{HH: We need to define p values corresponding to the data shown in plots. }
\red{second step: the p-values under $H_0$ are uniform.}

@@

Consider a linear regression model 
\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.

@@

A lineup including this test statistic is shown in Figure \ref{fig:test_category}. The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma^2})$. The test statistic, the plot containing the observed data, is randomly placed among these null plots. If the test statistic is identifiable the null hypothesis is rejected. % with a $p$-value of at most 0.05.

@@

\begin{table*}[hbtp] 
\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} ~~... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{0.5cm}m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
\hline\hline 
Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

@@

Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
\begin{eqnarray}\label{dist_p}
F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
  &=&\left\{ \begin{array}{ll}
          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
       \end{array} \right.     
\end{eqnarray}

Thus the density of $p_B$ is Uniform(0,1)  under $H_0$. As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.  In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

Now assume that individuals pick the plot that has the smallest $p$-value in the lineup plot. This leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as

@@

Figure \ref{fig:power_expected} shows the power of UMP test and expected power of visual test obtained from equation \eqref{power_exp}. Notice that the expected power of visual test is almost as good as the power of UMP test. In fact, the only loss we experience stems from comparing the test statistic to only a discrete number $m-1$ samples of the null distribution. Theorem \ref{thm:power} helps to compute the expected power of visual inference.

\begin{thm}\label{thm:power}
 Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
\end{thm}

{ \em proof}

Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
\begin{eqnarray*}
F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
  &=& \left[ x^{m-1}\right]^1_{1-t}\\
  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
\end{eqnarray*}

% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t

Thus expected power in equation \ref{power_exp} be
\begin{eqnarray*}
 Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
  &=& E(1-p_B)^{m-1}   \\
%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
\end{eqnarray*}

\begin{figure}[hbtp]
   \centering
       \scalebox{.65}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.}
       \label{fig:power_expected}
\end{figure}

@@

\subsection{Some Power Considerations}

Let us think of a lineup as a head-to-head comparison of the data plot and $m-1$ null plots. Let the data plot have a $p$ value of $p_B$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
We know that for each comparison the probability that the data plot has the smaller $p$ value is 
\begin{eqnarray*}
P(p_B < p_{0,i}) &=& 1 - P(p_{i,0} \le p_B) = 1- \int_0^1  P(p_{i,0} \le p_B \mid p_B=t) f_{p_B}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_B}(t) dt = 1 - \int_0^1 t f_{p_B}(t) dt = 1 - E[p_B],
\end{eqnarray*}
which, in particular, is independent of $p_{0,i}$ for all $i$.

Let us now define $X$ as the number of null plots in a lineup, for which the $p$ value $p_{0,i}$ is smaller than $p_B$ (i.e. under the assumption that the onlooker is able to identify the chart with the smallest $p$ value, they would pick any of the $X$  null plots over the data plot). 
Then $X \sim B_{m-1, E[p_B]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
\[
P(X=0) = \left(1 - E[p_B] \right)^{m-1} = P(p_B \le p_0),
\]
where $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$

\subsection{Expected number of choices}
Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
\[
E[X] = (m-1)(1-E[p_B]).
\]
With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
This should allow us to infer some of the signal strength $p_B$ as
\begin{equation} \label{plot_signal}
E[p_B] = 1 - E[X]/(m-1),
\end{equation}
i.e. by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.

Now assume that the same lineup plot is evaluated by $K$ independent observers. Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$

For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
\[
n_m\sim \left \{ 
\begin{array}{cl}
B_{K, (1-E[p_B])^{m-1}} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]
Figure \ref{fig:pval_power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e. reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$. The grey line shows probability of picking the data plot under $H_0$. We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$. (Mahbub, is that what we see in practice?) Yes, we see this pattern in our experiments as shown in figure \ref{fig:pval_pcorrect}}
   \label{fig:pval_power}
\end{figure}

>>>>

\setlength\parindent{0pt}

@@

\section{Introduction}

Any statistical analysis must include some statistical graphics. For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics. Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.

@@

\red{HH: We need to define p values corresponding to the data shown in plots. }

\blue{Let $D$ be a measure for the effect contained in a data set. We will assume that $D$ follow a distribution $F_D$. Let $p_D$ be the $p$-value associated with the data shown in the test statistic.

Similarly, we will assume that $p_{0,i}$, $i=1, ..., m-1$ are the $p$-values associated with data corresponding to the $m-1$ null plots. Since this data is generated consistently with the null hypothesis, the $p$-values follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$. }

\red{move power consideration up:}

Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
We know that for each comparison the probability that the data plot has the smaller $p$ value is 
\begin{eqnarray*}
P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
\end{eqnarray*}
which, in particular, is independent of $p_{0,i}$ for all $i$.

Let us now define $X$ as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$ (i.e. under the assumption that the onlooker is able to identify the chart with the smallest $p$-value, they would pick any of the $X$  null plots over the data plot). 
\red{HH: we need to make this assumption a bit more obvious - this is a crucial assumption, and we need to back it up later in the evaluation section.}

Then $X \sim B_{m-1, E[p_D]}$, and the probability that an onlooker will pick the data plot in a given lineup is 
\[
P(X=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
\]

%\red{pull the results for $P(p_B < p_{i,0})$ and $P(p_B < p_0)$ to the front, including the figure on $1/m$ versus power. These results are independent of the regression setting - they don't make any assumptions in terms of regression.}

Figure \ref{fig:pval_power} gives an overview of the probability of picking the data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. We see that for a $p$-value $p_D$ of about 0.15 for the data plot the signal in the plot is so weak, that we cannot distinguish the data plot from null plots in a lineup of size $m=20$. This pattern is also seen in our experiments as shown in figure \ref{fig:pval_pcorrect}.

Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{images/p_val_power_m.pdf} 
   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_D$. The grey line shows probability of picking the data plot under $H_0$. }
   \label{fig:pval_power}
\end{figure}

Now assume that the same lineup plot is evaluated by $K$ independent observers. W.l.o.g. let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, then $\sum_{k=1}^{m} n_k = K$, and $(n_1, n_2, ..., n_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

For the distribution of $n_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
we get
\[
n_m\sim \left \{ 
\begin{array}{ll}
B_{K, (1-E[p_D])^{m-1}} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]

Under $H_1$ we therefore get an estimate for the strength of the signal in the test statistic based on a lineup evaluation as 
\[
\widehat{E[p_D]} = 1- \left(\frac{n_m}{K}\right)^{1/(m-1)}.
\]

\red{We need comment on how this ties in to the rest of the paper: this is a theoretical consideration based on the assumption that observers are able to spot the plot with the smallest/a small $p$-value.}

@@

Consider a linear regression model 
\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.

@@

A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}. The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$. The test statistic, the plot containing the observed data, is randomly placed among these null plots. If the test statistic is identifiable the null hypothesis is rejected. % with a $p$-value of at most 0.05.

@@

\begin{table*}[hbtp] 
\caption{Test Statistics for Testing Hypothesis Related to Model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{0.5cm}m{3cm}m{2.5cm}m{2.6cm}m{5.5cm}} 
\hline\hline 
Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_bet_p.pdf}} \end{center} \end{minipage} & Residual vs $X_k$ plots. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plots of residuals & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.18}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.11}{\includegraphics{stat_interection.pdf}} \end{center} \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_nonlinear.pdf}} \end{center} \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.15}{\includegraphics{stat_sigma_box.pdf}} \end{center} \end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{2cm}\begin{center}  	\scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_goodness_simple.pdf}} \end{center} \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{2cm}\begin{center}  \scalebox{0.1}{\includegraphics{stat_intercept.pdf}} \end{center} \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

@@

\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
\begin{eqnarray}\label{dist_p}
F{p_B}(p) &=& Pr(p_B \le p)=1-Pr(1-p_B \le 1-p) \nonumber \\
  &=& 1-Pr(F_{|t|}^{-1}(1-p_B) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
  &=&\left\{ \begin{array}{ll}
          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
       \end{array} \right.     
\end{eqnarray}

Thus the density of $p_B$ is Uniform(0,1)  under $H_0$. As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{P_B}(p_B; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_B);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_B))}$$ derived from equation \ref{dist_p}  is a right skewed distribution.  
%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

Under the assumption that an observer is able to pick the plot with the smallest $p$-value from a lineup plot,  this leads to the decision to reject $H_0$ when $p_{B} < p_0$. Thus we have the expected power as

@@

Figure \ref{fig:power_expected} shows the power of the uniformly most powerful test (UMP) in comparison to the expected power of the visual test obtained from equation \eqref{power_exp}. Notice that the expected power of the visual test is almost as good as the power of UMP test. In fact, the only loss we experience stems from comparing the test statistic to only a discrete number of  $m-1$ samples from the null distribution.

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_B)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \ref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_B)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.65}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$.}
       \label{fig:power_expected}
\end{figure}

@@

%\subsection{Some Power Considerations}

%\subsection{Expected number of choices}
%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
%\[
%E[X] = (m-1)(1-E[p_B]).
%\]
%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
%This should allow us to infer some of the signal strength $p_B$ as
%\begin{equation} \label{plot_signal}
%E[p_B] = 1 - E[X]/(m-1),
%\end{equation}
%i.e. by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
%
